\documentclass{aiaa-tc}% insert '[draft]' option to show overfull boxes

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{calc}%allows for scaling figures with integer division/multiplication of existing lengths
\usepackage{wrapfig}
\usepackage{lipsum}
\usepackage[hidelinks]{hyperref}

\title{Design and Manufacture of an Open-Hardware 
 	University Rocket Airframe using Carbon Fiber}

\author{
Joseph Shields, Leslie Elwood, Erik Nelson, and Jacob East%, Brandon Bonner
	\thanks{Portland State University, Portland, OR 97201}
 }
% Data used by 'handcarry' option if invoked
\AIAApapernumber{2016}
\AIAAconference{Conference Name, Date, and Location}
\AIAAcopyright{\AIAAcopyrightD{2016}}
 
% Define commands to assure consistent treatment throughout document
\newcommand{\eqnref}[1]{(\ref{#1})}
\newcommand{\class}[1]{\texttt{#1}}
\newcommand{\package}[1]{\texttt{#1}}
\newcommand{\file}[1]{\texttt{#1}}
\newcommand{\BibTeX}{\textsc{Bib}\TeX}
\newcommand{\cots}{commercial off-the-shelf}
\newcommand{\weightReduction}{80\%}
\newcommand{\strengthIncrease}{??\%}

\begin{document}
\maketitle

\begin{abstract}
The amateur and university rocketry communities are rapidly reaching higher altitudes with more sophisticated rockets. However, most groups are still using heavy airframes made of metal or fiberglass. Commercial off-the-shelf airframes are either too expensive for low-budget university groups or too small to use as a platform for high altitude experiments. 
A capstone team of mechanical engineering seniors at Portland State University has developed a low-weight, modular carbon fiber airframe as an open-hardware technology for university rocketry. This team is continuing the work of a 2014 capstone team, who developed a carbon fiber layup process with promising results. 
This will enable low-budget groups like the Portland State Aerospace Society to explore high altitude science and compete in the university space race.  \end{abstract}
\section*{Nomenclature}
\begin{itemize}
	\item CFD, Computational Fluid Dynamics
	\item PSAS, Portland State Aerospace Society
	\item LV2, Launch Vehicle 2
	\item LV3, Launch Vehicle 3
\end{itemize}


\section{Introduction}
%REMEMBER TO SPECIFY ALL ACRONYMS/INITIALISMS!
The Portland State Aerospace Society (PSAS) is an interdisciplinary group of engineering students and alumni of Portland State University (PSU) with the long term goal of putting a cubesat into orbit with their own rocket. 
% info about cubesat market growth: http://www.spaceref.com/news/viewpr.html?pid=44940 
Their current airframe, named Launch Vehicle 2 (LV2), has served for over 12 years, representing 10 of the group's 13 launches, and hosted experiments ranging from custom patch antennas and long range WiFi technology to GPS navigation and a cold gas reaction control system (figure \ref{fig:L-12}). The LV2 platform is mostly constructed of aluminum with a fiberglass shell, with many of the parts having been fabricated in home garages. This makes for a robust but heavy design. Additionally, this airframe is built with a 4.5 inch inner diameter which PSAS's experiments have outgrown. 

The new airframe being designed, named Launch Vehicle 3 (LV3), aims to address these issues. The LV3 platform uses a 6 inch inner diameter, modules composed of carbon fiber and thin aluminum coupling rings, a carbon fiber nose cone, and a carbon fiber fin section. All of the airframe components connect via standardized rings, to accommodate future experimental modules and flight configurations.
The cylindrical LV3 airframe modules outperform the old design with an \weightReduction{} reduction in weight.
% Do we know what the increase in strength is? Also, can we confirm the weight reduction?

\section{Software}
\subsection{OpenRocket}
For the early design of amateur and university-scale rockets, \href{http://openrocket.sourceforge.net/}{OpenRocket} is very useful. 
It is an open-source Java application which simulates a rocket's flight. It provides a convenient interface for configuring the rocket, selecting commercial or custom parts, viewing the predictions, and performing some basic optimization. 

Using OpenRocket to guide the initial design allowed for rapid iteration of the fin geometry and the placement of non-airframe components. 

\subsection{OpenFOAM}
OpenFOAM is another open-source application. It provides a common interface for a wide range of CFD solvers.
Although it requires more up-front effort, it allows a more detailed perspective on what features of an airframe's geometry are the most critical. 
CFD analysis is probably not necessary for most amateur and university level airframe design, but this provides a good option for groups that can't afford a licence for a commercial CFD suite. 

OpenFoam was used to asses the heating at the tip of the nose, to determine if the epoxy matrix of the carbon fiber would degrade over repeated flights.

\subsection{Significance}
Many amateur and university rocketry groups use composite airframes. However, these designs use many layers of fabric, with low fiber volume fractions.
While easy and durable, this does not fully take advantage of their materials.



\section{Materials}

\subsection{Acquisition}
\section{Cylindrical Modules}

\subsection{Destructive Testing}

\section{Nose Cone}
To make the nose cone would take significant research into materials and ingenuity. The pre-impregnated carbon fiber material donated to us is a material with a 350 degree Fahrenheit cure cycle. Thus every element used henceforth must survive the same temperature. The initial manufacturing process involved a mandrel and the same procedure as the cylindrical modules. Machine Sciences agreed to donate the material and machine time to create this for PSAS however, lead time prevented them from delivering this to us in the time frame necessary for launch. 

After much consideration, it was decided that the next best method for creating the nose cone would be to use a mold. A 1 inch thick, high temp, 18 pound density machinable foam (FR 4718) was purchased from General Plastics and stacked to create a 4 inch thickness. ACP high-temperature resin epoxy is mixed with a 3.2:10 ratio was used to mix this resin before application and then ran through a heat cycle to cure. Next, the foam block was transported to ESCO to be run through a CNC Mill and returned. 

The mold experience mild delamination during milling and needed extra resin to hold together. Knowing that the machinable foam was porous and would likely absorb any fluid epoxy created by the high temperature environment, it was decided that the surface of the mold would be coated in the ACP resin and run through another heat cycle. 

\section{Fin Section}
\end{document}
